---
title: "children Excercise"
author: "Harish"
date: "02/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE}
suppressMessages(library(readxl))
suppressMessages(fit_database_exercise_normal <- read_excel("~/SharedFiles/ST606/2020/data/Exercise/fit_database_exercise_normal.xlsx"))
suppressMessages(fit_database_anthropometric_all<-read_excel("~/SharedFiles/ST606/2020/data/Exercise/fit_database_anthropometric_all.xlsx"))

```

```{r, include = FALSE}
str(fit_database_exercise_normal)
suppressWarnings(fit_database_exercise_normal$gender <- as.factor(fit_database_exercise_normal$gender))
suppressWarnings(fit_database_exercise_normal$`running distance (m)` <- as.numeric(fit_database_exercise_normal$`running distance (m)`))
suppressWarnings(fit_database_exercise_normal$`running time (s)` <- as.numeric(fit_database_exercise_normal$`running time (s)`))
suppressWarnings(fit_database_exercise_normal$`running speed (m/s)` <- as.numeric(fit_database_exercise_normal$`running speed (m/s)`))
suppressWarnings(fit_database_exercise_normal$`z-category (WHO)` <- as.factor(fit_database_exercise_normal$`z-category (WHO)`))
suppressWarnings(fit_database_exercise_normal$`pulse 0'` <- as.numeric(fit_database_exercise_normal$`pulse 0'`))
suppressWarnings(fit_database_exercise_normal$`pulse 1'` <- as.numeric(fit_database_exercise_normal$`pulse 1'`))
suppressWarnings(fit_database_exercise_normal$`pulse 5'` <- as.numeric(fit_database_exercise_normal$`pulse 5'`))
suppressWarnings(fit_database_exercise_normal$`pulse 10'` <- as.numeric(fit_database_exercise_normal$`pulse 10'`))
suppressWarnings(fit_database_exercise_normal$`SBP 0'` <- as.numeric(fit_database_exercise_normal$`SBP 0'`))
suppressWarnings(fit_database_exercise_normal$`SBP 1'` <- as.numeric(fit_database_exercise_normal$`SBP 1'`))
suppressWarnings(fit_database_exercise_normal$`SBP 5'` <- as.numeric(fit_database_exercise_normal$`SBP 5'`))
suppressWarnings(fit_database_exercise_normal$`SBP 10'` <- as.numeric(fit_database_exercise_normal$`SBP 10'`))
suppressWarnings(fit_database_exercise_normal$`DBP 0'` <- as.numeric(fit_database_exercise_normal$`DBP 0'`))
suppressWarnings(fit_database_exercise_normal$`DBP 1'`<- as.numeric(fit_database_exercise_normal$`DBP 1'`))
suppressWarnings(fit_database_exercise_normal$`DBP 5'`<- as.numeric(fit_database_exercise_normal$`DBP 5'`))
suppressWarnings(fit_database_exercise_normal$`DBP 10'`<- as.numeric(fit_database_exercise_normal$`DBP 10'`))
suppressWarnings(fit_database_exercise_normal$uniqueid <- row(fit_database_exercise_normal))
```

Q.(1) In this question we will discuss about the individuals who ran incorrect distance.

```{r}

suppressMessages(library(dplyr))
# no of observations with age less than 10.
lessthan10<- fit_database_exercise_normal %>% filter(fit_database_exercise_normal$`age bin` < 10)

# no of observations with age less than 10 but ran 2000.(2)
exc1 <- lessthan10 %>% filter(lessthan10$`running distance (m)` ==2000)

# no of observations with age greater than 10.
morethan10<-fit_database_exercise_normal %>% filter(fit_database_exercise_normal$`age bin` > 10)

exc2 <- morethan10 %>% filter(morethan10$`running distance (m)` ==1000)

incorrectboys <- exc2 %>% filter(exc2$gender=="boy")
incorrectgirls <- exc2 %>% filter(exc2$gender=="girl")
```

Important points

* 2 girls ran 2000 metres however, they were supposed to ran only 1000.
* 2251 observations found where running distance was only 1000 metres, but they were supposed to ran 2000.
* 922 Number of observations are found were boys ran incorrect distance. 
* 1329 Number of observations are found were girls ran incorrect distance. 
* So from the total observation of 2251 where individuals ran incorrect distance 59% were girls             observation.

```{r}
 suppressMessages(library(ggplot2))
 suppressMessages(library(ggpubr))
 suppressMessages(install.packages("ggpubr"))
 
 ggplot(exc2)+
geom_bar(mapping = aes(x= `age bin`, fill = gender))
  
ggplot(exc2)+
  geom_bar(mapping = aes(x= `age bin`))+
  facet_wrap(~gender)

```

* From the two plots we can say that boys and girls with age equal to 11 years were certainly more in numbers as compared to other age childrens who ran incorrect distance and with the exception of age bin 11 no of girls where 

```{r}
# To remove those observations were childrens ran incorrect distance. 
ID1 <- exc1$uniqueid
ID2 <- exc2$uniqueid
# normal dataset after removing wrong data for running distance.
cleandata <- fit_database_exercise_normal[-ID1,]
cleandata <- cleandata[-ID2,]
```

 Q.(2) Analysis of the observations were running distance is given but running time and running speed is missing. So assuming these are the observations were individuals left running track in between.
 
```{r}
# normal data set without Na values from running distance.
na_cases <- which(!complete.cases(cleandata$`running distance (m)`))
# na_cases is the dataset where running distance is NA.
naless <- cleandata[-na_cases,]
# naless is the dataset created after removinf NA values related to running distance.

na_cases1<- which(!complete.cases(naless$`running time (s)`))
# na_cases1 is the dataset where running time and running speed is NA.
incomplete <- naless[na_cases1,]
# # incomplete is the dataset where running time and running speed is NA in dataset naless where running distance is not NA or missing.

incompleteboys <- incomplete %>% filter(incomplete$gender =="boy")
incompletegirls <- incomplete %>% filter(incomplete$gender=="girl")
# From incomplete dataset, 3212 observations are of boys and 4087 observations are for girls, So 56% of individuals who may have left running in between are girls. 


```


```{r}

ggplot(incomplete)+
  geom_bar(mapping = aes(x= `age bin`))+
  facet_wrap(~gender)

ggplot(incomplete)+
  geom_bar(mapping = aes(x= `age bin`, fill = gender))

```

* From the plots we can se that for every age group, number of girls that may have left running in between are more than boys and age bin = 7, is the group where most number of incomplete data related to running is observed for both boys and girls.

```{r}
complete <-naless[-na_cases1,]
# complete is the data set created after removing NA values related to running distance, running speed and running       time from the dataset cleandata which was created after removing 2253 cases where the childrens ran incorrect          distance. 

remove_all <- which(!complete.cases(complete[14:25]))
cleanPSD <- complete[-remove_all,]
# clean PSD is the dataset which is obtained by removing all the missing value related to pulse rate, SBP and DBP.       Imputation is not performed here because missing values where more than 5%.
```

Q.(3) visualisation of individuals who had normal pulse rate, normal SBP and normal DBP before running or observations       which are taken at rest.                                                                                          
     normal pulse rate range :- 60-100
     normal SBP :- 90-120
     normal DBP :- 60-80
     
```{r}
col1 <- ifelse(cleanPSD$`pulse 0'` > 60 & cleanPSD$`pulse 0'` < 100 , "red","gree")
col2 <- ifelse(cleanPSD$`SBP 0'` > 90 & cleanPSD$`SBP 0'` < 120 , "red","gree")
col3 <- ifelse(cleanPSD$`DBP 0'` > 60 & cleanPSD$`DBP 0'` < 80 , "red","gree")

ggplot(cleanPSD)+
  geom_bar(mapping =aes(x= `pulse 0'`, col = col1), na.rm = TRUE)+
  facet_wrap(~`age bin` + `gender`)

ggplot(cleanPSD)+
  geom_bar(mapping =aes(x= `SBP 0'`, col = col2), na.rm = TRUE)+
  facet_wrap(~`age bin` + `gender`)

ggplot(cleanPSD)+
  geom_bar(mapping =aes(x= `DBP 0'`, col = col3), na.rm = TRUE)+
  facet_wrap(~`age bin` + `gender`)

```

* From the pulse rate plot it is clear that almost for every age group girls had high pulse rate as compared to          boys.The pulse rate for girls with age more than 11 was consistently more than 130 and even reaching to 140 for some   age groups which is very high.

* A normal resting pulse rate for adults ranges from 60 to 100 beats per minute. Generally, a lower pulse rate at rest   implies more efficient heart function and better cardiovascular fitness. from the plot, lower pulse rate observations   where more for boys as compared to girls and those observations were mainly observed after the age of 13.

* From the SBP plot it is clear that majority of the childrens with age ranging from 6-11 had normal SBP but from the    age of 12 the proportion that had high SBP increases and from the age of 12-18 the proportion with normal SBP was      less as compared to those that  had high SBP, with boys having large number of proportions with high SBP.

* From the plot of DBP it is clear that the children within age range from 6-10 had a big proportion of observations     with low DBP and a very small proportion of high DBP however from the age of 11 the observations with high DBP         increases. Both boys and girls nearly had the same proportion for out of range DBP.


* reference :- https://www.healthline.com/health/high-blood-pressure-hypertension/blood-pressure-reading-explained

* normal:- SBP(90-120) and DBP(60-80).
* Elevated:- SBP(120-130) and DBP(60-80).
* High blood pressure(stage 1) :- SBP(130-140) and DBP(80-89).
* High blood pressure(stage 2) :- SBP(140-179) and DBP(90-119).
* hypertensive crisis : SBP(>180) and DBP(>120).
* pulse rate : Active(40-60), normal(60-100), high(100-130), hypertensive crisis(130-180)

```{r}
cleanPSD$catsbp<- cut(cleanPSD$`SBP 0'`, breaks = c(60,90,120,130,140,180, 220), labels = c("low", "normal", "Elevated", "High S1", "High S2", "hypertensive crisis"), right = FALSE)

cleanPSD$catdbp <- cut(cleanPSD$`DBP 0'`, breaks = c(30,60,80,90,120,160),labels = c("low", "normal", "High S1", "High S2", "hypertensive crisis"), right = FALSE)

cleanPSD$catpulse <- cut(cleanPSD$`pulse 0'`, breaks = c(40,60,100,130,200),labels = c("active", "normal", "High", "hypertensive crisis"), right = FALSE)

```

* These are the plots to observe the proportions of each category for pulse, SBP, DBP.

```{r}

SS<-ggplot(data = cleanPSD) +
geom_bar(mapping = aes(fill = catsbp, x=""))

SBS <-ggplot(cleanPSD ,mapping = aes(x = catsbp, y = `SBP 0'` , fill = catsbp )) +
geom_boxplot()

DS <- ggplot(data = cleanPSD) +
geom_bar(mapping = aes(fill = catdbp, x=""))


DBS<- ggplot(cleanPSD ,mapping = aes(x = catdbp, y = `DBP 0'` , fill = catdbp )) +
geom_boxplot()

PS<- ggplot(data = cleanPSD) +
geom_bar(mapping = aes(fill = catpulse, x=""))

PBS<-ggplot(cleanPSD ,mapping = aes(x = catpulse, y = `pulse 0'` , fill = catpulse )) +
geom_boxplot()


fig <- ggarrange(SS, SBS,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1)


fig1 <- ggarrange( DS, DBS,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1)


fig2<- ggarrange( PS,PBS,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1)
fig


fig1


fig2

```

Q.(4) Analysis on the basis of individuals with most observations. (at rest)
```{r}
suppressMessages(library(tidyverse))
cleanPSD$ID <- as.factor(cleanPSD$ID)
length(levels(cleanPSD$ID))
invisible(sort(table(cleanPSD$ID), decreasing = TRUE))
ID1441 <- filter(cleanPSD, ID==1441)
ID360 <- filter(cleanPSD, ID==360)

pulsetarget <- 100
SBPtarget <- 120
DBPtarget <- 80
```

* ID 1441 have 20 observations and ID 360 have 18 so we one use these observations for the analysis of one individual.

ID-1441
```{r}

P<- ggplot(data= ID1441,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")

Q<-ggplot(data= ID1441,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
  geom_hline(yintercept=SBPtarget, color="red")

R<-ggplot(data= ID1441,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
  geom_hline(yintercept=DBPtarget, color="red")

figure7 <- ggarrange(
  P,                
  ggarrange(Q, R, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  labels = "A"  
  ) 
  
figure7
```

ID-360
```{r}

X <- ggplot(data= ID360,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")

Y <- ggplot(data= ID360,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
geom_hline(yintercept=SBPtarget, color="red")

Z <- ggplot(data= ID360,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
geom_hline(yintercept=DBPtarget, color="red")

figure6 <- ggarrange(
  X,                
  ggarrange(Y, Z, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  labels = "A"  
  ) 
  
figure6

```

*  This is done to to observe pattern where each ID have 17 observation.
```{r}
 ID67 <- filter(cleanPSD, ID== 67 )
 ID81 <- filter(cleanPSD, ID== 81)
 ID299 <- filter(cleanPSD, ID== 299)
 ID367 <- filter(cleanPSD, ID== 367)
 ID2583 <- filter(cleanPSD, ID== 2583)
 ID3608 <- filter(cleanPSD, ID== 3608)
 ID4546 <- filter(cleanPSD, ID== 4546)
 obs_17 <- rbind(ID67,ID81,ID299,ID367,ID2583,ID3608,ID4546)
```


```{r}
P17 <- ggplot(data= obs_17,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")+
  facet_wrap(~ID)

S17 <- ggplot(data= obs_17,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
geom_hline(yintercept=SBPtarget, color="red")+
  facet_wrap(~ID)

D17 <- ggplot(data= obs_17,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
geom_hline(yintercept=DBPtarget, color="red")+
  facet_wrap(~ID)

figure5 <- ggarrange(
  P17,                
  ggarrange(S17, D17, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  labels = "A"  
  ) 
  
figure5
```

*  This is done to to observe pattern where each ID have 16 observation.
```{r}
ID72 <- filter(cleanPSD, ID==72)
ID147 <- filter(cleanPSD, ID==147)
ID227 <- filter(cleanPSD, ID==227)
ID278 <- filter(cleanPSD, ID==278)
ID291 <- filter(cleanPSD, ID==291)
ID307 <- filter(cleanPSD, ID==307)
ID322 <- filter(cleanPSD, ID== 322)
ID334 <- filter(cleanPSD, ID==334)
ID352 <- filter(cleanPSD, ID==352)
ID388 <- filter(cleanPSD, ID==388)
ID405 <- filter(cleanPSD, ID==405)
ID920 <- filter(cleanPSD, ID==920)
ID931 <- filter(cleanPSD, ID==931)
ID1321 <- filter(cleanPSD, ID==1321)
ID3612 <- filter(cleanPSD, ID==3612)
ID3668 <- filter(cleanPSD, ID==3668)
ID4771 <- filter(cleanPSD, ID==4771)

obs_16 <- rbind(ID72, ID147, ID227, ID278, ID291, ID307, ID322, ID334, ID352, ID388, ID405, ID920, ID931, ID1321, ID3612, ID3668, ID4771)

```


```{r}
P16 <- ggplot(data= obs_16,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")+
  facet_wrap(~ID)

S16 <-ggplot(data= obs_16,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
geom_hline(yintercept=SBPtarget, color="red")+
  facet_wrap(~ID)

D16 <-ggplot(data= obs_16,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
geom_hline(yintercept=DBPtarget, color="red")+
  facet_wrap(~ID)

figure4 <- ggarrange(
  P16,                
  ggarrange(S16, D16, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  labels = "A"  
  ) 
  
figure4

```

* This is done to check whether there is any correlation pulse, SBP and DBP observations with respect to time.

```{r}
suppressMessages(library(GGally))
suppressWarnings(ggpairs(cleanPSD, columns= 14:17,aes(col=gender), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= 18:21,aes(col=gender), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= 22:25,aes(col=gender),  progress = FALSE))

```

* From the plot for pulse, SBP and DBP highest correlation is between variables at 5 min and variables at 10, so    plotting pulse, SBP and DBP at 5 VS pulse, SBP and DBP at 10. 

```{r}

p <- ggplot(cleanPSD,aes(`pulse 5'`,`pulse 10'`))
q <- ggplot(cleanPSD,aes(`SBP 5'`,`SBP 10'`))
r <- ggplot(cleanPSD,aes(`DBP 5'`,`DBP 10'`))

 P5_10 <-p + stat_bin2d()+
scale_fill_gradient(low = "lightblue1", high = "steelblue4")

S5_10 <-q + stat_binhex()+
scale_fill_gradient(low = "lightblue1", high = "steelblue4")

D5_10 <- r + stat_density2d(aes(fill = ..level..), geom = "polygon")

figure3 <- ggarrange(
  P5_10,                
  ggarrange(S5_10, D5_10, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  labels = "A"  
  ) 
  
figure3

```


Q.(5) Check whether there is any relation between pulse, SBP and DBP.
```{r}
suppressMessages(library(GGally))
suppressWarnings(ggpairs(cleanPSD, columns= c(14,18,22), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= c(15,19,23), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= c(16,20,24), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= c(17,21,25), progress = FALSE))

```

* From the plots, no relation is clear between any variable and from the correlation values we can say pulse has no relation with either SBP and DBP.However, there is almost 40-60 percent correlation between SBP and DBP for all observations.

Q.(6)  Observe the average pattern of pulse, SBP and DBP observations with respect to time over the period of experimental years.

```{r}

pulse0 <- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(pulse0Mean = mean(`pulse 0'`)) %>%
ggplot(aes(x= `measurement date`, y=pulse0Mean)) + geom_line()

pulse1 <- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(pulse1Mean = mean(`pulse 1'`)) %>%
ggplot(aes(x= `measurement date`, y=pulse1Mean)) + geom_line()

pulse5 <- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(pulse5Mean = mean(`pulse 5'`)) %>%
ggplot(aes(x= `measurement date`, y=pulse5Mean)) + geom_line()

pulse10 <- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(pulse10Mean = mean(`pulse 10'`)) %>%
ggplot(aes(x= `measurement date`, y=pulse10Mean)) + geom_line()

figure2 <- ggarrange(pulse0, pulse1, pulse5,pulse10,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
figure2
```


* From the 1 plot we can say the average resting pulse throughout the experimental years was under 100 and hence normal.

* From the 2 plot we can say the average pulse just after the run from 2008-2013 was consistently over 135 which is      very high but after 2013 the average pulse rate decreases and even getting less than 120 is 2016 which can be          considered good as pulse 1 is taken just after completing the run.

* From the 3 plot we can say the average pulse after 5 minutes of the run was consistently less than 116 and reaches     its lowest average in 2018.

* From the 4 plot we can say the average pulse after 10  minutes of the run was nearly same to the pulse rate after 5    minutes as they both satayed around than 110-115. The lowest average for pulse10 was recorded in 2008.

```{r}
SBP0 <- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(SBP0Mean = mean(`SBP 0'`)) %>%
ggplot(aes(x= `measurement date`, y=SBP0Mean)) + geom_line()

SBP1 <-cleanPSD%>%
group_by(`measurement date`) %>%
summarise(SBP1Mean = mean(`SBP 1'`)) %>%
ggplot(aes(x= `measurement date`, y=SBP1Mean)) + geom_line()

SBP5 <-cleanPSD%>%
group_by(`measurement date`) %>%
summarise(SBP5Mean = mean(`SBP 5'`)) %>%
ggplot(aes(x= `measurement date`, y=SBP5Mean)) + geom_line()

SBP10 <-cleanPSD%>%
group_by(`measurement date`) %>%
summarise(SBP10Mean = mean(`SBP 10'`)) %>%
ggplot(aes(x= `measurement date`, y=SBP10Mean)) + geom_line()

figure1 <- ggarrange(SBP0, SBP1, SBP5,SBP10,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
figure1

```

* Average SBP at rest after 2009 remained under 120 and hence normal but in 2008 it was around 126 which is considered high at rest.

* Average SBP after 1 min of the run was between 155-125 which is above normal but this value decreases every year from it's peak in 2008 to it's lowest in 2016.

* Average SBP after 5 min of the run was between 130-105. This value decreases every year from it's peak in 2008 to it's lowest in 2015. After 2009 it's value was consistently less than 120 which is normal. So, we can say after 5 min of the run average SBP tends to get normal. 

* After 10 mins of run SBP was normal throughout the experimental years.

 
```{r}

DBP0<- cleanPSD %>%
group_by(`measurement date`) %>%
summarise(DBP0Mean = mean(`DBP 0'`)) %>%
ggplot(aes(x= `measurement date`, y=DBP0Mean)) + geom_line()

DBP1<- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(DBP1Mean = mean(`DBP 1'`)) %>%
ggplot(aes(x= `measurement date`, y=DBP1Mean)) + geom_line()

DBP5<- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(DBP5Mean = mean(`DBP 5'`)) %>%
ggplot(aes(x= `measurement date`, y=DBP5Mean)) + geom_line()

DBP10<- cleanPSD%>%
group_by(`measurement date`) %>%
summarise(DBP10Mean = mean(`DBP 10'`)) %>%
ggplot(aes(x= `measurement date`, y=DBP10Mean)) + geom_line()

figure <- ggarrange(DBP0, DBP1, DBP5,DBP10,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
figure
```

* Average distolic blood pressure throughout the experimental years at rest, after 1 min , after 5min and after 10 min   of the run remained under the normal category and reaching it's lowest value in 2018 and that to was normal so we can   say running exercise doesn't seem to have much impact on the distolic blood pressure.


Q.(7) Analysis on the basis of  age, running distance, pulse rate, SBP and DBP after completing the specified run respectively.

```{r}
age10 <- cleanPSD %>% filter(cleanPSD$`age bin` == 10)
exc3 <- age10 %>% filter(age10$`running distance (m)` == 1000)
exc4 <- age10 %>% filter(age10$`running distance (m)` == 2000)

dist1000<- cleanPSD %>% filter(cleanPSD$`age bin` <10)
dist2000 <- cleanPSD %>% filter(cleanPSD$`age bin` >10)
cdist1000 <- rbind(dist1000, exc3)
cdist2000 <- rbind(dist2000, exc4)

```

* Age10 is the dataset were the age bin of childrens is 10 and the some of the individuals of age bin 10   ran 1000      metres while some ran 2000 metres, So I created two seperate datasets exc3 for distance equals   and exc4 where        distance equals 2000. Then I merge exc3 with dist1000 which is the dataset where running distance was 1000 and exc4    with dist2000 where the running distance was 2000.

```{r}
```









