---
title: "children Excercise"
author: "Harish"
date: "02/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
suppressMessages(library(readxl))
suppressMessages(fit_database_exercise_normal <- read_excel("~/SharedFiles/ST606/2020/data/Exercise/fit_database_exercise_normal.xlsx"))
suppressMessages(fit_database_anthropometric_all<-read_excel("~/SharedFiles/ST606/2020/data/Exercise/fit_database_anthropometric_all.xlsx"))

```

```{r}
str(fit_database_exercise_normal)
suppressWarnings(fit_database_exercise_normal$gender <- as.factor(fit_database_exercise_normal$gender))
suppressWarnings(fit_database_exercise_normal$`running distance (m)` <- as.numeric(fit_database_exercise_normal$`running distance (m)`))
suppressWarnings(fit_database_exercise_normal$`running time (s)` <- as.numeric(fit_database_exercise_normal$`running time (s)`))
suppressWarnings(fit_database_exercise_normal$`running speed (m/s)` <- as.numeric(fit_database_exercise_normal$`running speed (m/s)`))
suppressWarnings(fit_database_exercise_normal$`z-category (WHO)` <- as.factor(fit_database_exercise_normal$`z-category (WHO)`))
suppressWarnings(fit_database_exercise_normal$`pulse 0'` <- as.numeric(fit_database_exercise_normal$`pulse 0'`))
suppressWarnings(fit_database_exercise_normal$`pulse 1'` <- as.numeric(fit_database_exercise_normal$`pulse 1'`))
suppressWarnings(fit_database_exercise_normal$`pulse 5'` <- as.numeric(fit_database_exercise_normal$`pulse 5'`))
suppressWarnings(fit_database_exercise_normal$`pulse 10'` <- as.numeric(fit_database_exercise_normal$`pulse 10'`))
suppressWarnings(fit_database_exercise_normal$`SBP 0'` <- as.numeric(fit_database_exercise_normal$`SBP 0'`))
suppressWarnings(fit_database_exercise_normal$`SBP 1'` <- as.numeric(fit_database_exercise_normal$`SBP 1'`))
suppressWarnings(fit_database_exercise_normal$`SBP 5'` <- as.numeric(fit_database_exercise_normal$`SBP 5'`))
suppressWarnings(fit_database_exercise_normal$`SBP 10'` <- as.numeric(fit_database_exercise_normal$`SBP 10'`))
suppressWarnings(fit_database_exercise_normal$`DBP 0'` <- as.numeric(fit_database_exercise_normal$`DBP 0'`))
suppressWarnings(fit_database_exercise_normal$`DBP 1'`<- as.numeric(fit_database_exercise_normal$`DBP 1'`))
suppressWarnings(fit_database_exercise_normal$`DBP 5'`<- as.numeric(fit_database_exercise_normal$`DBP 5'`))
suppressWarnings(fit_database_exercise_normal$`DBP 10'`<- as.numeric(fit_database_exercise_normal$`DBP 10'`))
suppressWarnings(fit_database_exercise_normal$uniqueid <- row(fit_database_exercise_normal))
```

Q.(1) In this question we will discuss about the individuals who ran incorrect distance.

```{r}

suppressMessages(library(dplyr))
# no of observations with age less than 10.
lessthan10<- fit_database_exercise_normal %>% filter(fit_database_exercise_normal$`age bin` < 10)

# no of observations with age less than 10 but ran 2000.(2)
exc1 <- lessthan10 %>% filter(lessthan10$`running distance (m)` ==2000)
# 2 girls ran 2000 metres however, they were supposed to ran only 1000.

# no of observations with age greater than 10.
morethan10<-fit_database_exercise_normal %>% filter(fit_database_exercise_normal$`age bin` > 10)
# 2251 observations found where running distance was only 1000 metres, but they were supposed to ran 2000.
exc2 <- morethan10 %>% filter(morethan10$`running distance (m)` ==1000)

incorrectboys <- exc2 %>% filter(exc2$gender=="boy")
# 922 Number of observations are found were boys ran incorrect distance. 
incorrectgirls <- exc2 %>% filter(exc2$gender=="girl")
# 1329 Number of observations are found were girls ran incorrect distance. 
# So from the total observation of 2251 where individuals ran incorrect distance 59% were girls observation.
 
 suppressMessages(library(ggplot2))
 ggplot(exc2)+
  geom_bar(mapping = aes(x= `age bin`, fill = gender))
  
 ggplot(exc2)+
  geom_bar(mapping = aes(x= `age bin`))+
  facet_wrap(~gender)
# From the two plots we can say that boys and girls with age equal to 11 years were certainly more in numbers as compared to other age childrens who ran incorrect distance.


# To remove those observations were childrens ran incorrect distance. 
ID1 <- exc1$uniqueid
ID2 <- exc2$uniqueid
# normal dataset after removing wrong data for running distance.
cleandata <- fit_database_exercise_normal[-ID1,]
cleandata <- cleandata[-ID2,]

```
 Q.(2) Analysis of the observations were running distance is given but running time and running speed is missing. So assuming these are the observations were individuals left running track in between.
 
```{r}
# normal data set without Na values from running distance.
na_cases <- which(!complete.cases(cleandata$`running distance (m)`))
# na_cases is the dataset where running distance is NA.
naless <- cleandata[-na_cases,]
# naless is the dataset created after removinf NA values related to running distance.

na_cases1<- which(!complete.cases(naless$`running time (s)`))
# na_cases1 is the dataset where running time and running speed is NA.
incomplete <- naless[na_cases1,]
# # incomplete is the dataset where running time and running speed is NA in dataset naless where running distance is not NA or missing.

incompleteboys <- incomplete %>% filter(incomplete$gender =="boy")
incompletegirls <- incomplete %>% filter(incomplete$gender=="girl")
# From incomplete dataset, 3212 observations are of boys and 4087 observations are for girls, So 56% of individuals who may have left running in between are girls. 

ggplot(incomplete)+
  geom_bar(mapping = aes(x= `age bin`))+
  facet_wrap(~gender)

ggplot(incomplete)+
  geom_bar(mapping = aes(x= `age bin`, fill = gender))

# From the plots we can se that for every age group, number of girls that may have left running in between are more than boys and age bin = 7, is the group where most number of incomplete data related to running is observed for both boys and girls.
```


```{r}
complete <-naless[-na_cases1,]
# complete is the data set created after removing NA values related to running distance, running speed and running      time from the dataset cleandata which was created after removing 2253 cases where the childrens ran incorrect         distance. 

remove_all <- which(!complete.cases(complete[14:25]))
cleanPSD <- complete[-remove_all,]
# clean PSD is the dataset which is obtained by removing all the missing value related to pulse rate, SBP and DBP. Imputation is not performed here because missing values where more than 5%.
```

Q.(3) visualisation of individuals who had normal pulse rate, normal SBP and normal DBP before running or observations which are taken at rest.
     normal pulse rate range :- 60-100
     normal SBP :- 90-120
     normal DBP :- 60-80
```{r}
col1 <- ifelse(cleanPSD$`pulse 0'` > 60 & cleanPSD$`pulse 0'` < 100 , "red","gree")
col2 <- ifelse(cleanPSD$`SBP 0'` > 90 & cleanPSD$`SBP 0'` < 120 , "red","gree")
col3 <- ifelse(cleanPSD$`DBP 0'` > 60 & cleanPSD$`DBP 0'` < 80 , "red","gree")

ggplot(cleanPSD)+
  geom_bar(mapping =aes(x= `pulse 0'`, col = col1), na.rm = TRUE)+
  facet_wrap(~`age bin` + `gender`)

ggplot(cleanPSD)+
  geom_bar(mapping =aes(x= `SBP 0'`, col = col2), na.rm = TRUE)+
  facet_wrap(~`age bin` + `gender`)

ggplot(cleanPSD)+
  geom_bar(mapping =aes(x= `DBP 0'`, col = col3), na.rm = TRUE)+
  facet_wrap(~`age bin` + `gender`)
# From the pulse rate plot it is clear that almost for every age group girls had high pulse rate as compared to boys.The pulse rate for girls with age more than 11 was consistently more than 130 and even reaching to 140 for some age groups which is very high.

# A normal resting pulse rate for adults ranges from 60 to 100 beats per minute. Generally, a lower pulse rate at rest implies more efficient heart function and better cardiovascular fitness. from the plot, lower pulse rate observations where more for boys as compared to girls and those observations were mainly observed after the age of 13.

# From the SBP plot it is clear that majority of the childrens with age ranging from 6-11 had normal SBP but from the age of 12 the proportion that had high SBP increases and from the age of 12-18 the proportion with normal SBP was less as compared to those that  had high SBP, with boys having large number of proportions with high SBP.

# From the plot of DBP it is clear that the children within age range from 6-10 had a big proportion of observations with low DBP and a very small proportion of high DBP however from the age of 11 the observations with high DBP increases. Both boys and girls nearly had the same proportion for out of range DBP.
```

```{r}

# reference :- https://www.healthline.com/health/high-blood-pressure-hypertension/blood-pressure-reading-explained
# normal:- SBP(90-120) and DBP(60-80).
# Elevated:- SBP(120-130) and DBP(60-80).
# High blood pressure(stage 1) :- SBP(130-140) and DBP(80-89).
# High blood pressure(stage 2) :- SBP(140-179) and DBP(90-119).
# hypertensive crisis : SBP(>180) and DBP(>120).
# pulse rate : Active(40-60), normal(60-100), high(100-130), hypertensive crisis(130-180)
cleanPSD$catsbp<- cut(cleanPSD$`SBP 0'`, breaks = c(60,90,120,130,140,180, 220), labels = c("low", "normal", "Elevated", "High S1", "High S2", "hypertensive crisis"), right = FALSE)

cleanPSD$catdbp <- cut(cleanPSD$`DBP 0'`, breaks = c(30,60,80,90,120,160),labels = c("low", "normal", "High S1", "High S2", "hypertensive crisis"), right = FALSE)

cleanPSD$catpulse <- cut(cleanPSD$`pulse 0'`, breaks = c(40,60,100,130,200),labels = c("active", "normal", "High", "hypertensive crisis"), right = FALSE)

ggplot(data = cleanPSD) +
geom_bar(mapping = aes(fill = catsbp, x=""))

ggplot(data = cleanPSD) +
geom_bar(mapping = aes(fill = catdbp, x=""))

ggplot(data = cleanPSD) +
geom_bar(mapping = aes(fill = catpulse, x=""))

ggplot(cleanPSD ,mapping = aes(x = catsbp, y = `SBP 0'` , fill = catsbp )) +
geom_boxplot()

ggplot(cleanPSD ,mapping = aes(x = catdbp, y = `DBP 0'` , fill = catdbp )) +
geom_boxplot()

ggplot(cleanPSD ,mapping = aes(x = catpulse, y = `pulse 0'` , fill = catpulse )) +
geom_boxplot()

```


Q.(4) Analysis on the basis of individuals with most observations. ( at rest)

```{r}
suppressMessages(library(tidyverse))
cleanPSD$ID <- as.factor(cleanPSD$ID)
length(levels(cleanPSD$ID))
invisible(sort(table(cleanPSD$ID), decreasing = TRUE))
# ID 1441 have 20 observations and ID 360 have 18 so we one use these observations for the analysis of one individual.
ID1441 <- filter(cleanPSD, ID==1441)
ID360 <- filter(cleanPSD, ID==360)

pulsetarget <- 100
SBPtarget <- 120
DBPtarget <- 80

ggplot(data= ID1441,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")

ggplot(data= ID1441,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
  geom_hline(yintercept=SBPtarget, color="red")

ggplot(data= ID1441,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
  geom_hline(yintercept=DBPtarget, color="red")


ggplot(data= ID360,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")

ggplot(data= ID360,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
geom_hline(yintercept=SBPtarget, color="red")

ggplot(data= ID360,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
geom_hline(yintercept=DBPtarget, color="red")

# This is done to to observe pattern where each ID have 17 observation.
ID67 <- filter(cleanPSD, ID== 67 )
ID81 <- filter(cleanPSD, ID== 81)
ID299 <- filter(cleanPSD, ID== 299)
ID367 <- filter(cleanPSD, ID== 367)
ID2583 <- filter(cleanPSD, ID== 2583)
ID3608 <- filter(cleanPSD, ID== 3608)
ID4546 <- filter(cleanPSD, ID== 4546)
obs_17 <- rbind(ID67,ID81,ID299,ID367,ID2583,ID3608,ID4546)

ggplot(data= obs_17,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")+
  facet_wrap(~ID)

ggplot(data= obs_17,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
geom_hline(yintercept=SBPtarget, color="red")+
  facet_wrap(~ID)

ggplot(data= obs_17,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
geom_hline(yintercept=DBPtarget, color="red")+
  facet_wrap(~ID)
 
ID72 <- filter(cleanPSD, ID==72)
ID147 <- filter(cleanPSD, ID==147)
ID227 <- filter(cleanPSD, ID==227)
ID278 <- filter(cleanPSD, ID==278)
ID291 <- filter(cleanPSD, ID==291)
ID307 <- filter(cleanPSD, ID==307)
ID322 <- filter(cleanPSD, ID== 322)
ID334 <- filter(cleanPSD, ID==334)
ID352 <- filter(cleanPSD, ID==352)
ID388 <- filter(cleanPSD, ID==388)
ID405 <- filter(cleanPSD, ID==405)
ID920 <- filter(cleanPSD, ID==920)
ID931 <- filter(cleanPSD, ID==931)
ID1321 <- filter(cleanPSD, ID==1321)
ID3612 <- filter(cleanPSD, ID==3612)
ID3668 <- filter(cleanPSD, ID==3668)
ID4771 <- filter(cleanPSD, ID==4771)

obs_16 <- rbind(ID72, ID147, ID227, ID278, ID291, ID307, ID322, ID334, ID352, ID388, ID405, ID920, ID931, ID1321, ID3612, ID3668, ID4771)


ggplot(data= obs_16,
aes(x=`measurement date`, y=`pulse 0'`)) + geom_line()+
geom_hline(yintercept=pulsetarget, color="red")+
  facet_wrap(~ID)

ggplot(data= obs_16,
aes(x=`measurement date`, y=`SBP 0'`)) + geom_line()+
geom_hline(yintercept=SBPtarget, color="red")+
  facet_wrap(~ID)

ggplot(data= obs_16,
aes(x=`measurement date`, y=`DBP 0'`)) + geom_line()+
geom_hline(yintercept=DBPtarget, color="red")+
  facet_wrap(~ID)
```

```{r}
suppressMessages(library(GGally))
suppressWarnings(ggpairs(cleanPSD, columns= 14:17,aes(col=gender), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= 18:21,aes(col=gender), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= 22:25,aes(col=gender),  progress = FALSE))

```


```{r}
# From the plot for pulse, SBP and DBP highest correlation is between pulse at 5 and pulse at 10.
p <- ggplot(cleanPSD,aes(`pulse 5'`,`pulse 10'`))
q <- ggplot(cleanPSD,aes(`SBP 5'`,`SBP 10'`))
r <- ggplot(cleanPSD,aes(`DBP 5'`,`DBP 10'`))

p + stat_bin2d()+
scale_fill_gradient(low = "lightblue1", high = "steelblue4")

q + stat_binhex()+
scale_fill_gradient(low = "lightblue1", high = "steelblue4")

r + stat_density2d(aes(fill = ..level..), geom = "polygon")
```


Q.(5) Analysis on the basis of  age, running distance, pulse rate, SBP and DBP after completing the specified run respectively.

```{r}
age10 <- cleanPSD %>% filter(cleanPSD$`age bin` == 10)
exc3 <- age10 %>% filter(age10$`running distance (m)` == 1000)
exc4 <- age10 %>% filter(age10$`running distance (m)` == 2000)

dist1000<- cleanPSD %>% filter(cleanPSD$`age bin` <10)
dist2000 <- cleanPSD %>% filter(cleanPSD$`age bin` >10)
cdist1000 <- rbind(dist1000, exc3)
cdist2000 <- rbind(dist2000, exc4)
# Age10 is the dataset were the age bin of childrens is 10 and the some of the individuals of age bin 10 ran 1000 metres while some ran 2000 metres, So I created two seperate datasets exc3 for distance equals 1000 and exc4 where distance equals 2000. Then I merge exc3 with dist1000 which is the dataset where running distance was 1000 and exc4 with dist2000 where the running distance was 2000.


```

```{r}
# This is to check whether there is any relation between pulse, SBP and DBP.
suppressMessages(library(GGally))
suppressWarnings(ggpairs(cleanPSD, columns= c(14,18,22), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= c(15,19,23), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= c(16,20,24), progress = FALSE))

suppressWarnings(ggpairs(cleanPSD, columns= c(17,21,25), progress = FALSE))


# From the plots, no relation is clear between any variable and from the correlation values we can say pulse has no relation with either SBP and DBP.However, there is almost 40-60 percent correlation between SBP and DBP for all observations.

```



```{r}

```


```{r}

```

 
```{r}
```
```{r}
```


```{r}
```
```{r}
```


```{r}
```
```{r}
```


```{r}
```
```{r}
```


```{r}
```
```{r}
```


```{r}
```
```{r}
```


```{r}
```
```{r}
```


```{r}
```
```{r}
```


```{r}
```














